---
title: "04_prepare_data_to_generate_derived_dataset_file"
author: "Mireia Valle"
date: "2024-01-12"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warming = FALSE)
```

# Description

Script to remove outliers from the species occurrence data based on distance.

# Load libraries ---------------------------------------------------
```{r}
library(here) #enable easy file referencing 
library(ggplot2) #plotting
library(tidyverse)
library(R.utils)#loadToEnv
library(glue)#glue
```

#Loading species id
```{r}
species <- read.csv (here::here ("data","derived_data", "02_species_occurrences","species_list_FINAL_hemispheres_automatic.csv"),  sep = ",")

species <- species %>% 
  dplyr:: select (-X) 

n <- data.frame(species$SP)

names(n)

n <- n %>%
  dplyr::rename (id = "species.SP")

n <- str_remove(n$id, "SP")

n <- as.numeric (n)

```

#Load species datasets
```{r}

for (i in n){

#Load species presence data without outliers
  
SP <- loadToEnv(file.path("data", "derived_data" , "02_species_occurrences", "02_species_occurrences_no_outliers_automatic", paste0("SP",i, ".RData")))[[paste0("SP",i)]]
  

derived_dataset <- SP %>% 
    group_by(datasetKey, source) %>% 
    summarize (n = n())
    
write.csv(derived_dataset, file=here::here ("data", "derived_data" , "02_species_occurrences", "04_derived_datasets" , paste0("derived_dataset_gbif_obis_SP", i, "_", SP$scientificName[i],".csv")), row.names=F)
  
}

```

# Generate an overall derived dataset for all downloaded occurrences
```{r}

files <- list.files(here::here ("data", "derived_data" , "02_species_occurrences", "04_derived_datasets"), pattern = "*.csv", full.names = TRUE)

data <- lapply(files, read.csv, header = TRUE, stringsAsFactors = FALSE, row.names = NULL)

df <- do.call(rbind.data.frame, data)

df2 <- df %>% 
  group_by(datasetKey, source) %>% 
  summarize (n = sum(n)) %>% 
  drop_na()

write.csv(df2, file=here::here ("data", "derived_data" , "02_species_occurrences", "04_derived_datasets" , "derived_dataset_obis_gbif.csv"), row.names=F)

df3 <- df %>% 
  dplyr::filter(source == "gbif.org") %>% 
  group_by(datasetKey) %>% 
  summarize (n = sum(n)) %>% 
  drop_na()

write.csv(df3, file=here::here ("data", "derived_data" , "02_species_occurrences", "04_derived_datasets" , "derived_dataset_gbif.csv"), row.names=F)

```

With this derived dataset information file created you have to make sure to deposit your dataset somewhere publicly accessible (e.g. Zenodo). When you’re ready, go to the derived dataset tool from GBIF: https://www.gbif.org/derived-dataset/register (you might need to register if you are not registered yet) fill in the form, upload the last .csv we have created (derived_dataset_gbif.csv) and submit the form. You will get a doi for your derived dataset.



